---
title: "modelr notes"
author: "Steven Aguilar"
date: "February 21, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(modelr)
library(purrr)
library(lubridate)
```

## Fundamentals of Models

Models can show us useful information about our model. We will look at both patterns and residuals. 
  
Split data into:
  
* training set (about 60%) - can do anything with this you want as many times as you want
* query set (about 20%) - compare models or visualizations by hand but not as part of an automated process
* test set (about 20%) - use only once to test your final model after you have selected it

We will define a family of models that express a precise, but generic, pattern that we want to identify. Then, we will generate a fitted model by finding the model from the family that best fits the data.

We begin by examing a simple dataset: sim1 that is part of modelr. 
  
```{r}
sim1
ggplot(sim1, aes(x,y))+
  geom_point()


```
Looks linear. So let's think about fitting a line to the data. This will be our family of models... y =a1+a2x. so let's set up a tible that contains lots of lines with slopes between -5 and 5 and intercepts between -20 and 40. We will then plot these lines on top of our points

```{r}
models <- tibble(
  a1 = runif(250, -20, 40),
  a2 = runif(250, -5, 5)
)

ggplot(sim1, aes(x,y))+
  geom_abline(
    aes(intercept = a1, slope = a2),
    data=models, alpha = 1/4)+
  geom_point()


```

```{r}
#Compute the predicted y

model1 <- function(a, data){
  a[1] + data$x * a[2]
}

model1(c(7,2), sim1)

#Compute the distance using the same inputs
measure_distance <- function(mod, data){
  diff <- data$y - model1(mod, data)
  sqrt(mean(diff^2))
}

measure_distance(c(7,2), sim1)
```

```{r}
# measure_distance expectrs a numeric vector of length 2 followed by the dataset name

sim1_dist <- function(a1,a2){
  measure_distance(c(a1,a2), sim1)
}

models <- models %>%
  mutate(dist = map2_dbl(a1, a2, sim1_dist))
models

#plot best ten models, colored by distance
ggplot(sim1, aes(x, y))+
  geom_point(size = 2 , color = "grey30") +
  geom_abline(
    aes(intercept = a1, slope = a2, color = -dist),
    data = filter(models, rank(dist) <= 10)
  )

```

We could instead look at the coordinates versus the distance

```{r}
models %>%
  ggplot(aes(a1,a2)) +
  geom_point(
    data = filter(models, rank(dist) <= 10),
    size = 4, color = "red") +
  geom_point(aes(color = -dist))
```

```{r}
grid <- expand.grid(
  a1 = seq(-10,10, length = 50),
  a2 = seq(1,4, length = 50)
) %>%
  mutate(dist = map2_dbl(a1, a2, sim1_dist))
head(grid %>% arrange(dist))

sim1 %>%
  ggplot(aes(x, y))+
  geom_point(size = 2 , color = "grey30") +
  geom_abline(
    aes(intercept = a1, slope = a2, color = -dist),
    data = filter(grid, rank(dist) <= 10)
  )
```
  
```{r}
best <- optim(c(4,2), measure_distance, data = sim1)
best$par

sim1 %>%
  ggplot(aes(x,y)) +
  geom_point(size=2, color="grey30")+
  geom_abline(
    aes(intercept = best$par[1], slope = best$par[2])
  )
```

## Visualizing models
```{r}
grid <- sim1 %>%
  data_grid(x)
grid

grid2 <- mtcars %>%
  data_grid(disp, wt)
grid2
tail(grid2)

sim1_mod <- lm(y ~ x, data=sim1)

grid <- grid %>%
  add_predictions(sim1_mod)

sim1 %>%
  ggplot(aes(x)) +
  geom_point(aes(y=y)) +
  geom_point(aes(y = pred), data = grid,
             color = "red",
             size = 3)

```
   
```{r}
sim1 <- sim1 %>%
  add_residuals(sim1_mod)

sim1 %>%
  ggplot(aes(x, resid))+
  geom_ref_line(h=0) + geom_point()


```
  
```{r}
#Categorical variable
sim2

sim2 %>%
  ggplot(aes(x,y)) + geom_point()

mod2 <- lm(y ~ x, data = sim2)
grid <- sim2 %>%
  data_grid(x)%>%
  add_predictions(mod2)
grid

sim2 %>%
  ggplot(aes(x=x)) +
  geom_point(aes(y = y)) +
  geom_point(
    data = grid,
    aes(y = pred),
    color = "red",
    size = 3
  )

```

## two predictors
####PAY ATTENTION
```{r}
ggplot(sim3, aes(x1, y)) +
  geom_point(aes(color=x2))

mod1 <- lm(y ~ x1 + x2, data = sim3)
mod2 <- lm(y ~ x1 *x2, data = sim3)

grid <- sim3 %>%
  data_grid(x1,x2) %>%
  gather_predictions(mod1, mod2)
grid

sim3 %>%
  ggplot(aes(x1, y, color = x2))+
  geom_point(aes(color = x2))+
  geom_line().............
```

```{r}
sim3 <- sim3 %>%
  gather_residuals(mod1, mod2)

sim3 %>%
  ggplot(aes(x1, resid, color = x2))+
  geom_point()+
  facet_grid(model ~ x2)
```

## two continuous variables
```{r}
mod1 <- lm(y ~ x1 + x2, data = sim4)
mod2 <- lm(y ~ x1 * x2, data = sim4)

grid <- sim4 %>%
  data_grid(
    x1 = seq_range(x1, 5),
    x2 = seq_range(x2, 5)
  ) %>%
  gather_predictions(mod1, mod2)
grid


grid %>%
  ggplot(aes(x1, x2)) +
  geom_tile()
```

