---
title: "Purrr"
author: "Steven Aguilar"
date: "February 14, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Iteration

imperative programming = for and while loops
function programming = each common for loop pattern gets its own function

### For loops

3 componenents:

*initialization of output (use vector (type, length))
*for (i in seq_along(df))
* body - code that does the work an dputs results into th eoutput you initialized

```{r}
library(tidyverse)

df <- tibble(
a = rnorm(20,4,1),
b = rnorm(20, 4, 1.2),
c = rnorm(20, 0, 1),
d = rnorm(20, -1, 1)
)

median(df$a)
median(df$b)
median(df$c)
median(df$d)

#instead use a for loop

output <- vector("double", ncol(df))

for (i in seq_along(output)) {
  output[i] <- median(df[[i]])
}

output
```

### For loop variations

* modifying an existing object
* looping over names or vlues instead of indices
* unknown output lenght
* unknown sequence length

```{r}
rescale01 <- function(x){
  rng <- range(x, na.rm = TRUE)
  (x-rng[1]) / (rng[2] - rng[1])
}

for (i in seq_along(df)) {
  df[[i]] <- rescale01(df[[i]])
}

df
```

#### Looping patterns

* for (x in xs) - usually for side effects like plotting or saving a file
* for (nm in names(xs)) - helps to use the name in a plot title or filename


##### Unknown Output Length

Do NOT grow the output with each iteration. This is very inefficient and will up the computer time. Let's look at simulating random vectors of random lenghts

```{r}
means <- c(0,1,2)

output <- double()

for(i in seq_along(means)){
  n <- sample(100,1)
  output <- c(output, rnorm(n, means[[i]]))
}

str(output)

##better approach
out <- vector("list", length(means))

for (i in seq_along(means)){
  n <- sample(100,1)
  out[[i]] <- rnorm(n, means[[i]])
}

str(out)

str(unlist(out))
```

other similar situations:

* generating a long string. If you are using paste() to continue growin gth estring with each iteraion, instead save the output ...
as a character vector, then combine into a single string using *paste(output, collapse="")*

* if you are using rbind() in each iteration, save the ...



while loop

keep looping until a certain condition occurs
Dangerous because you could get stuck in an infinite loop if the condition never occurs. I personaly prefer a for loop with a maximum number of iterations.


# Homework or classwork: p. 321 () #1, #3

A)
```{r}
output <- vector("double", ncol(mtcars))
for (i in seq_along(output)) {
  output[i] <- mean(mtcars[[i]])
}

output

```

B)
```{r}
output <- vector("character", ncol(nycflights13::flights))
for (i in seq_along(output)) {
  output <- str(nycflights13::flights[[i]])
}
output
```

C)
```{r}
output <- vector("double", ncol(iris))
for (i in seq_along(output)){
  output[i] <- length(unique(iris[[i]]))
}

output
```

D)
```{r}
df <- c(-10,0,10,100)
output <- vector("list", length(df))
for (i in seq_along(df)){
  output[[i]] <- rnorm(10, df[[i]])
}
output
```

###Functional Programming

```{r}
df <- tibble(
a = rnorm(20,4,1),
b = rnorm(20, 4, 1.2),
c = rnorm(20, 0, 1),
d = rnorm(20, -1, 1)
)

#for loop
output <- vector("double", ncol(df))

for (i in seq_along(output)) {
  output[i] <- median(df[[i]])
}
output

#put into a function
col_median <- function(df){
  output <- vector("double", ncol(df))
  for (i in seq_along(df)) {
    output[[i]] <- median(df[[i]])
  }
  output
}

col_median(df)

#Computing IQR
col_IQR <- function(df){
  output <- vector("double", ncol(df))
  for (i in seq_along(df)) {
    output[[i]] <- IQR(df[[i]])
  }
  output
}

col_IQR(df)



```

#####Passing Functions
```{r}
col_info <- function(df, func){
  output <- vector("double", ncol(df))
  for (i in seq_along(df)) {
    output[[i]] <- func(df[[i]])
  }
  output
}

col_info(df, median)
col_info(df, mean)

```

First we will figure out how to solve a problem for a single element of a list. Then use purrr functions to generalize our solution to every element in the list.

#### map functions
Take a vector as input, applies a function to each piece and returns a new vector with the same names and length as th einput vecotr. Pick your function based on the type of output you want.

* map() makes a list
* map_lgl makes a logical vector
* map_int() makes an integer vector
* map_dbl() makes a double vector
* map_chr() makes a character vector


```{r}
map_dbl(df, median)
map_dbl(df, IQR)

df %>%
  map_dbl(median)

df%>%
  map_dbl(IQR)

# map also preserves the names associated with the vector
map_dbl(df, median, trim = 0.5)
```


#### Shortcuts in map functions that save typing

```{r}
mtcars %>% split(.$cyl)
models <- mtcars %>%
  split(.$cyl) %>%
  map(function(df) lm(mpg~wt, data = df))
models

#Shortcut for writing the same exact code
models <- mtcars %>%
  split(.$cyl) %>%
  map(~lm(mpg~wt, data=.)) #here . stands for the current list element
models

# getting the R Squared for each separate regression
models %>%
  map(summary)%>%
  map_dbl(~.$r.squared)

names(summary(models[[1]]))

models %>%
  map(summary)%>%
  map("fstatistic")%>%
  map_dbl("value")

```

#### Dealing with failures

* safely() returns a list of 2 elements, one of which is null

  -- first element is result if not error, NULL if error
  -- second element is error message if there is an error, NULL if no error
  
* possibly() always successful; provide....
* quietly() instead of capturing errors, it captures printed output, message, and warnings.


#### Mapping over Multiple Arguements

* map2 for iterating over two different arguements
* pmap for 

```{r}
mu <- list(-2,3,15)
sigma <- list(0.5,1,2)
map2(mu, sigma, rnorm, n=10) %>% str()

n <- list(5,10,15)
args1 <- list(n =n, mean=mu, sd=sigma)
args1 %>%
  pmap(rnorm) %>%
  str()

args2 <- tibble(
  n = c(5,10,15),
  mean = c(-2,3,15),
  sd = c(0.5,1,2)
)

args2 %>% 
  pmap(rnorm) %>%
  str()


```

#### Invoke Map
```{r}
# put all the functions in a vector or list

f <- list("runif", "rnorm", "rpois")

param <- list(
  list(min = -1, max = 1),
  list(sd = 2),
  list(lambda = 4)
)

invoke_map(f, param, n=5)



args3 <- tribble(
  ~f,     ~params,
  "runif", list(min = -1, max=1),
  "rnorm", list(sd=2),
  "rpois", list(lambda = 4)
)

temp <- args3 %>%
  mutate(args3 = invoke_map(f, param, n=5))

temp$args3 %>% str()
```























